Latent Predicate Networks: Concept Learning with Probabilistic Context-Sensitive Grammars

### 2014-10-20 drafts/cfp/draft.tex
Changes to draft.

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{listings}
\usepackage{anyfontsize}
\usepackage{subcaption}
\usepackage{graphicx}
\title{Latent Predicate Networks: Concept Learning with Probabilistic Context-Sensitive Grammars}
\author{Eyal Dechter,  Josh Rule, Josh Tenenbaum}

\begin{document}
\maketitle

@@

Range Concatenation Grammars (RCGs) are a class of string grammars
that represent all and only those languages that can be parsed in time
polynomial in the length of the target
string~\cite{boullier2005range}. An RCG $G=(N, T, V, P, S)$ is a
5-tuple where $N$ is a finite set of predicate symbols, $T$ is a set
of terminal symbols, $V$ is a set of variable symbols, P is a finite
set of $M \geq 0$ clauses of the form $\psi_0 \rightarrow \psi_1 \dots
\psi_M$, and $S \in N$ is the \emph{axiom}. Each $\psi_m$ is a term of
the form $A(\alpha_1, \dots, \alpha_{\mathcal{A}(A)}$, where $A \in
N$, $\mathcal{A}(A)$ is the arity of $A$, and each $\alpha_i \in (T
\cup V)^*$ is an argument of $\psi_m$. We call the left hand side term
of any clause the \emph{head} of that clause and its predicate symbol
is the \emph{head predicate}.

@@

We extend RCGs to Probabilistic Range Concatenation Grammars (PRCGs)
by annotating each clause $C_k \in P$ with probabilities $p_k$ such
that for all predicates ${A \in N$ $\sum_{k:head(C_k)=A} p_k = 1}$. A
PRCGs defines a distribution over strings $x$ by sampling from
derivations of $S(x)$ according to the product of probabilities of
clauses used in that derivation. This is a well defined distribution
as long as no probability mass is placed on derivations of infinite
length; in this paper, we only consider PRCGs with derivations of
finite length, so we need not worry about this requirement.

@@

Our goal is to understand how children can obtain semantic knowledge
about a system of concepts from the structure of relavent perceptual
stimuli. How does a child learn that prefixing ``great-'' corresponds
to adding an additional generation in kinship relationships? How does
a child learn that ``just before'' is after ``a while before'' which
is before ``a while after''? How does a child learn that if
Figure~\ref{pairOfApples} is ``a pair of apples'' then
Figure~\ref{singleApple} is a single apple?

Our proposal is that some interesting fragments of human conceptual
knowledge can be represented as structured relations between
linguistic tokens and other perceptual inputs; that the rules that
define these relations implicitly define the concepts that are
learned; and that these rules can be learned using the tools of
hierarchical Bayesian inference.

For example, the concept TWENTY SEVEN is defined as the collection of
representations ``27'', ``twenty seven'', ``3 times 9,'' along with a
system of symbolic rules that translate among these representations
and facilitate answered various queries. According to our hypothesis,
the difference between TWENTY EIGHT and TWENTY SEVEN is not that they
are represented in the mind by distinct mental symbols but rather that
the rules that operate on their various representations treat them in
systematically different ways.

This view implies that an attempt to construct a computational account
of concept learning in humans needs to ground out that learning in the
relavent perceptual stimuli.

Contribution:
\begin{itemize}
\item We present Generate Rewrite Systems (GRS) -- a formalism that borrows
  heavily from Linear Context-Free Rewriting Systems (CITE)
\item We show how GRS, though restricted in representational capactiy,
  can capture substantial cognitively plausible first-order conceptual knowledge 
\item We propose a heirarchical Bayesian model for GRS that provides a coherent framework in which a GRS can be learned given evidence
\item We show how GRS can be transformed into PRISM programs and learned within that probabilstic logic programming system
\item We provide computational experiments that provide evidence that
  our model can provide an account of some kinds of complex conceptual
  knowledge can be inferred from noisy and sparse data
\end{itemize}

@@

\section{The Model}

\begin{figure}[t]
		\includegraphics[width=\linewidth]{lpn/lpn.pdf}
		\caption{A schematic LPN where $m = |L|,\, n = |N|-1$,\, arity$(A_i) = 2$, and arity$(S) = 1$.}
		\label{fig:schematic}
\end{figure}

Now we can describe the model we introduce here, the Latent Concept
Grammar (LCG). We introduce the LCG as a machine learning model to
learn a distribution over instances of relations over strings. Unlike
a PRCG, which places a distribution over terms RCGs consist of a set
of rewrite rules, called clauses, each of which takes a termwhich
operate on a set of terms (predicates applied to arguments) to produce
new terms. Each rewrite rule is of the form:

Suppose we are given some positive instances $y_1, \dots,
y_M$ of relations $R_1, \dots, R_N$. Our model assumes that each
positive instance $y_m=R_n(\vec(x_m))$ is sampled i.i.d from a latent
stochastic grammar (LSG), $\mathcal{K}$. $\mathcal{K}$ consists of
some fixed number $K$ of clauses (i.e. rules) $C_1, \dots, C_K$ over
the $N$ observable relations plus a fixed number of $H$ hidden
relations $T_1, \dots, T_H$. The \emph{architecture} of the model is
the exact specification of the parameters $K$ and $H$ and the clauses
$\mathcal{C}$. To each clause $C_i$ is associated a positive real
weights $\theta_i$.

The clauses of $\mathcal{K}$ are Horn clauses in a first order logic
with string concatenation. We will use PROLOG notation, in which lower
case letters are predicates and constants and uppercase letters are
variables. If two symbols are adjacent it means that the corresponding
strings are concatenated. For example, the following would is a valid
clause: $a(XY, XV) \leftarrow b(X, UV), a(X, X)$. In this rule $a$ and
$b$ are relations with arity two.

Let $C_1, \dots, C_{N(R)} = C(R)$ be the set of clauses whose head is
relation $R$ and let $\Theta_R$ be the set of corresponding
weights. Let $p_R(C_n) \sim \text{Discrete}(\Theta_R)$; that is, $p_R$
is a distribution over the clauses headed by $R$, such that each
clauses is sampled in proportion to its weight parameter.

A positive instance of this relation is sampled from $\mathcal{K}$ by
arbitrarily (i.e. uniformly at random) choosing a relation $R$ and
instantiating that relation by sampling a clause from $p_R$. Recursing
on this procedure for every atom in the body of this clause binds the
variables in the clause's head.

Given this assumed generative process for the evidence, learning a
program involves inferring the parameter $\theta$ given a prior
probability distribution over their values. We place an independent
Dirichlet distribution over the parameters of every relation. This is
a natural choice that enables us to promote sparse programs, i.e.,
programs in which most of the mass is placed on a small number of
predicates and rules.

>>>>

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{listings}
\usepackage{anyfontsize}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{titling}

\setlength{\droptitle}{-10em}   % This is your set screw

\title{Latent Predicate Networks: Concept Learning with Probabilistic Context-Sensitive Grammars}
\author{Eyal Dechter,  Josh Rule, Josh Tenenbaum}

\begin{document}
\vspace{-10cm}
\maketitle

@@

\subsection{Probabilistic Range Concatentation Grammars}
Range Concatenation Grammars (RCGs) are a class of string grammars
that represent all and only those languages that can be parsed in time
polynomial in the length of the target
string~\cite{boullier2005range}. An RCG $G=(N, T, V, P, S)$ is a
5-tuple where $N$ is a finite set of predicate symbols, $T$ is a set
of terminal symbols, $V$ is a set of variable symbols, P is a finite
set of $M \geq 0$ clauses of the form $\psi_0 \rightarrow \psi_1 \dots
\psi_M$, and $S \in N$ is the \emph{axiom}. Each $\psi_m$ is a term of
the form $A(\alpha_1, \dots, \alpha_{\mathcal{A}(A)}$, where $A \in
N$, $\mathcal{A}(A)$ is the arity of $A$, and each $\alpha_i \in (T
\cup V)^*$ is an argument of $\psi_m$. We call the left hand side term
of any clause the \emph{head} of that clause and its predicate symbol
is the \emph{head predicate}.

@@

We extend RCGs to Probabilistic Range Concatenation Grammars (PRCGs)
by annotating each clause $C_k \in P$ with probabilities $p_k$ such
that for all predicates ${A \in N, \, \sum_{k:head(C_k)=A} p_k = 1}$. A
PRCGs defines a distribution over strings $x$ by sampling from
derivations of $S(x)$ according to the product of probabilities of
clauses used in that derivation. This is a well defined distribution
as long as no probability mass is placed on derivations of infinite
length; in this paper, we only consider PRCGs with derivations of
finite length, so we need not worry about this requirement.

\subsection{Architecture}

An LPN is an PRCG with $K$ binary predicates $A_1, \dots, A_K$ with a
densely connected structure. The architecture of the LPNs used in this
work is described in Figure~\ref{fig:architecture}. Specifically, for
each latent predicate $A_k$,

\begin{figure}[t]
		\includegraphics[width=\linewidth]{lpn/lpn.pdf}
		\caption{A schematic LPN where $m = |L|,\, n = |N|-1$,\, arity$(A_i) = 2$, and arity$(S) = 1$.}
		\label{fig:architecture}
\end{figure}

% Now we can describe the model we introduce here, the Latent Concept
% Grammar (LCG). We introduce the LCG as a machine learning model to
% learn a distribution over instances of relations over strings. Unlike
% a PRCG, which places a distribution over terms RCGs consist of a set
% of rewrite rules, called clauses, each of which takes a termwhich
% operate on a set of terms (predicates applied to arguments) to produce
% new terms. Each rewrite rule is of the form:

@@

% Our goal is to understand how children can obtain semantic knowledge
% about a system of concepts from the structure of relavent perceptual
% stimuli. How does a child learn that prefixing ``great-'' corresponds
% to adding an additional generation in kinship relationships? How does
% a child learn that ``just before'' is after ``a while before'' which
% is before ``a while after''? How does a child learn that if
% Figure~\ref{pairOfApples} is ``a pair of apples'' then
% Figure~\ref{singleApple} is a single apple?

% Our proposal is that some interesting fragments of human conceptual
% knowledge can be represented as structured relations between
% linguistic tokens and other perceptual inputs; that the rules that
% define these relations implicitly define the concepts that are
% learned; and that these rules can be learned using the tools of
% hierarchical Bayesian inference.

% For example, the concept TWENTY SEVEN is defined as the collection of
% representations ``27'', ``twenty seven'', ``3 times 9,'' along with a
% system of symbolic rules that translate among these representations
% and facilitate answered various queries. According to our hypothesis,
% the difference between TWENTY EIGHT and TWENTY SEVEN is not that they
% are represented in the mind by distinct mental symbols but rather that
% the rules that operate on their various representations treat them in
% systematically different ways.

% This view implies that an attempt to construct a computational account
% of concept learning in humans needs to ground out that learning in the
% relavent perceptual stimuli.

% Contribution:
% \begin{itemize}
% \item We present Generate Rewrite Systems (GRS) -- a formalism that borrows
%   heavily from Linear Context-Free Rewriting Systems (CITE)
% \item We show how GRS, though restricted in representational capactiy,
%   can capture substantial cognitively plausible first-order conceptual knowledge 
% \item We propose a heirarchical Bayesian model for GRS that provides a coherent framework in which a GRS can be learned given evidence
% \item We show how GRS can be transformed into PRISM programs and learned within that probabilstic logic programming system
% \item We provide computational experiments that provide evidence that
%   our model can provide an account of some kinds of complex conceptual
%   knowledge can be inferred from noisy and sparse data
% \end{itemize}

@@

% \section{The Model}

% Suppose we are given some positive instances $y_1, \dots,
% y_M$ of relations $R_1, \dots, R_N$. Our model assumes that each
% positive instance $y_m=R_n(\vec(x_m))$ is sampled i.i.d from a latent
% stochastic grammar (LSG), $\mathcal{K}$. $\mathcal{K}$ consists of
% some fixed number $K$ of clauses (i.e. rules) $C_1, \dots, C_K$ over
% the $N$ observable relations plus a fixed number of $H$ hidden
% relations $T_1, \dots, T_H$. The \emph{architecture} of the model is
% the exact specification of the parameters $K$ and $H$ and the clauses
% $\mathcal{C}$. To each clause $C_i$ is associated a positive real
% weights $\theta_i$.

% The clauses of $\mathcal{K}$ are Horn clauses in a first order logic
% with string concatenation. We will use PROLOG notation, in which lower
% case letters are predicates and constants and uppercase letters are
% variables. If two symbols are adjacent it means that the corresponding
% strings are concatenated. For example, the following would is a valid
% clause: $a(XY, XV) \leftarrow b(X, UV), a(X, X)$. In this rule $a$ and
% $b$ are relations with arity two.

% Let $C_1, \dots, C_{N(R)} = C(R)$ be the set of clauses whose head is
% relation $R$ and let $\Theta_R$ be the set of corresponding
% weights. Let $p_R(C_n) \sim \text{Discrete}(\Theta_R)$; that is, $p_R$
% is a distribution over the clauses headed by $R$, such that each
% clauses is sampled in proportion to its weight parameter.

% A positive instance of this relation is sampled from $\mathcal{K}$ by
% arbitrarily (i.e. uniformly at random) choosing a relation $R$ and
% instantiating that relation by sampling a clause from $p_R$. Recursing
% on this procedure for every atom in the body of this clause binds the
% variables in the clause's head.

% Given this assumed generative process for the evidence, learning a
% program involves inferring the parameter $\theta$ given a prior
% probability distribution over their values. We place an independent
% Dirichlet distribution over the parameters of every relation. This is
% a natural choice that enables us to promote sparse programs, i.e.,
% programs in which most of the mass is placed on a small number of
% predicates and rules.
